---
title: "Introduction"
subtitle: "Analysis of Animal Movement Data with R 2 (together with Brian Smith)"
format: 
  clean-revealjs:
    progress: true
    code-line-numbers: false
author:
  - name: Johannes Signer
    orcid: 0000-0002-1771-7775
    email: jmsigner@gmail.com
    affiliations: University of Göttingen, Germany & Freelancer
date: last-modified
execute:
  echo: true
  warning: false
  message: false
filters: 
  - timer
revealjs-plugins:
  - codefocus
---

```{r setup}
#| include: false
library(tidyverse)

knitr::opts_knit$set(
  root.dir = here::here("01 Introduction/")
)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.height = 4.5,
  fig.width = 4.5
)
```


# Welcome

## To get started

- Welcome to this online course on animal movement. 
- We are Brian and Johannes.
- Who are you? *Pair up and introduce your partner*

------

### Ask your partner: 

::::: {.coluns}

:::: {.column width = "60%"}

- What is your background?
- Where do you come from?
- Where do you study/work?
- What is your study organism?
- Why do you attend this course?
  
::::

:::: {.column width = "40%"}
  
:::{.timer #timer-1 seconds=180 starton=interaction}
:::

::::
  
:::::
  
## Outline of the course

### Day 1
1. Introduction and recap (J)
2. HSF/RSFs (B)
3. SSFs (J)

### Day 2
4. SSF 2 (B)
5. Interpretation (J)
6. Mixed RSF/SSF and multiple models (J)


------

### Day 3
7. GAM-SSF (B)
8. Validation (B)
9. HMM (J)


### Day 4
10. State-switching SSFs (J)
11. Simulating from fitted SSFs (J)
12. Q&A (both)


**Throughout the whole course and also for the exercises, if possible, please use your own data. **

--- 

## Some logistics

- Four days (Monday to Thursday) from 14:00 to 18:00 Berlin time.
- We split these 4-hour block into 3 - 4 chunks, each roughly structured like this: 
  - Lecture ~ 30 min
  - R walkthrough ~ 20 min
  - Introduction of exercises ~ 5 min
- A 20 min break approx. from 15:50 to 16:10.

------
    
- Lectures will be held via zoom.
- During the whole workshop we have a slack channel where you ask questions (we will monitor the channel during the course, feel free to ask questions there also outside the course hours).

- All course material is available here: <https://github.com/jmsigner/movement_ecology_2>

::: {.notes}
Say that you can download the whole repo in one go or clone it. 
:::

## Packages that we will use

- We will mainly use the `amt` package, but also occasionally other packages for movement analysis. 

- See also `required_packages.R` for a list of all packages that we need and to get the latest version of all packages. 

# Some conventions in R

- `%>%` or `|>`: Pipes the output of one function to a next function. We will discuss this further later on. 
- `::` to access a namespace form a package. 
- `data.frame` or `tibble`?

::: {.fragment}

- We often use `here::here("path to a file")`, when reading in a file. 
- The first `here` calls the function `here()` from the package `here`. 
- The function `here()` dynamically creates the absolute path to the project root. 

```{r}
here::here()
```

:::: {.notes}
This means, even though our computers differ, relative to the project root, we all have the same directory structure. 
::::

:::

# Movement data (in R)

## What we assume you are familiar with

- importing (movement) data to R ([see here](https://github.com/jmsigner/movement_ecology_1/tree/main/04%20Random%20Walks))
- familiarity with some packages (e.g., `amt`)
- familiarity with random walks and correlated random walks ([see here](https://github.com/jmsigner/movement_ecology_1/tree/main/04%20Random%20Walks))
- with home ranges ([see here](https://github.com/jmsigner/movement_ecology_1/tree/main/06%20Home%20Ranges))
- how to work with multiple instances ([see here](https://github.com/jmsigner/movement_ecology_1/tree/main/07%20Multiple%20Instances))

## Equal sampling rate

- Many methods that we deal with assume equal sample rates (i.e., the time distance between consecutive relocation is the same). 
- In `amt` the function `track_resample()` allows to do this. 


## Next: Code walkthrough

Briefly recap/illustrate: 

1. How to create a track
2. How to resample a track
3. Calculating an minimum conve polygon (MCP) home range
4. How to deal with multiple animals/instances
