---
title: "GAM-SSF Exercise"
author: "Brian J. Smith"
date: "22 April 2026"
output: 
  html_document:
    theme: default
    css: "../css/exercises.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

Let's apply what we've learned about GAM-SSFs. Refer to the lecture slides (<a href="07a_lecture.html" target="_blank" rel="noopener noreferrer">07a_lecture.html</a>) and the walkthrough code (`07b_walkthrough.R`) as a refresher.

We encourage you to use your own data for this exercise.

If you don't have your own data, you can use example data built into `amt`. You can access the fisher data and some covariate rasters like this:

```{r amt_fisher, eval = FALSE}
# Load `amt`
library(amt)

# Location data as 'track_xyt'
dat <- amt_fisher

# Habitat data as list
hab <- get_amt_fisher_covars()

```

## Instructions

*Warning: complex GAMs can take a long time to fit. If this is your first attempt, consider starting with a simple model!*

1. Load your data or the example data. Subset to a single individual.

    * Recall that iSSFs assume a constant step duration. Make sure your data are (within a reasonable tolerance) regularly sampled before you continue.
    * You will want a fairly large sample size to accomodate a wiggly function.
    
2. Fit an iSSF using a GAM with at least one smooth term. Choose from:

    * Non-linear selection.
    * Time-varying selection.
    * Spatial random effects.

3. If you'd like, you can also include random slopes and/or hierarchical smooths.
    
4. Make at least one plot showing the RSS (or log-RSS) for one of your habitat variables.

    * Recall, $RSS(x_1, x_2) = w(x_1)/w(x_2)$.
    * What does this plot tell you about the biology of your system?
    

<br>
<hr>
<br>


